Methods for Large Scale Learning to Rank

### 2014-09-29 report/Chapters/CrossBenchmarkComparison.tex
nog een nDCG gevonden
crossbenchmark comparison verbeterd vanuit IPM versie

In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis. To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used. This Normalised Winning Number takes only those data sets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all data sets on which it has been evaluated. We will redefine the indicator function $I$ in order to only take into account those data sets that an algorithm is evaluated on, as \\

@@

From now on this adjusted version of Winning Number will be references to as \emph{Normalised Winning Number}. The mathematical definition of Normalised Winning Number is\\

@@

\noindent
where Ideal Winning Number is defined as\\

@@

\section{Results \& Discussion}
The following subsections provide the performance of Learning to Rank methods in terms of Normalised Winning Number for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}. Performance of the Learning to Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of data sets on which the method has been evaluated on the horizontal axis. The further to the right, the more certain we can be about the performance of the Learning to Rank method. The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number data sets evaluated, are identified as the best performing methods and are labelled with the name of the method.

\subsection{NDCG@3}
Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{nDCG}@3 metric. Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw Normalised Winning Number data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\

@@

LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2. LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}). FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\

@@

LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0. LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of data sets, but score decreasingly well in terms of Normalised Winning Number. These results are highly in agreement with the \ac{NDCG}@3 comparison. The only modification compared to the \ac{NDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{NDCG}@5 comparison, but not in the \ac{NDCG}@3 comparison. Like in the \ac{NDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.

@@

LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{NDCG}@10 comparison. For LambdaNeuralRank these results are again based on AOL dataset measurements. LambdaMART showed the highest \ac{NDCG}@10 performance for the MSLR-WEB10k dataset. The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}. SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices. DCMP is not in the set in contrast with the \ac{NDCG}@3 and \ac{NDCG}@5 comparison.

\subsection{MAP}
Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning to Rank methods for the \ac{MAP} metric. Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\

@@

Where comparisons on the \ac{NDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results. RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{NDCG}-comparisons. In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\

LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons. Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison. A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number. SmoothRank performed a Normalised Winning Number of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods. RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations. No reported \ac{NDCG} performance was found in the literature study for RE-QR. There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning to Rank methods.

\subsection{Cross-metric}
Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}. Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.\\

@@

LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number. LRUF is the only method that achieved this in all \ac{NDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison. With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{NDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{NDCG} performance as opposed to the \ac{MAP} performance. This was to be expected, because the cross-metric comparison input data of three \ac{NDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\

@@

C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalised Winning Number measures, but have low Ideal Winning Number measures. Further evaluation runs of these methods on benchmark data sets that they are not yet evaluated on are desirable. The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more data sets than the number of data sets that we included evaluation results for in this meta-analysis. Some of the DirectRank measurements could not be used because measurements on some data sets were only available in graphical form and not in raw data.\\

@@

\section{Limitations}
In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark. By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set. It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some data sets than on other data sets. \\

@@

Another limitation lies in the meta-analysis methodology. Using evaluation results published by other researchers relies on the honesty of those researchers. Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results. Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely. By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.

\section{Conclusions}
We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
\begin{description}
\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
\end{description}
Although no closing arguments can be formulated on which Learning to Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\

@@

In the following chapters of this thesis, concerning parallel execution of the Learning to Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning to Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet. Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.

>>>>

In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis. To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used. This \ac{NWN} takes only those data sets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all data sets on which it has been evaluated. We will redefine the indicator function $I$ in order to only take into account those data sets that an algorithm is evaluated on, as \\

@@

From now on this adjusted version of Winning Number will be references to as \ac{NWN}. The mathematical definition of \ac{NWN} is\\

@@

\noindent
where \ac{IWN} is defined as\\

@@

\section{Results \& Discussion}
The following subsections provide the performance of Learning to Rank methods in terms of \ac{NWN} for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}. Performance of the Learning to Rank methods is plotted with \ac{NWN} on the vertical axis and the number of data sets on which the method has been evaluated on the horizontal axis. The further to the right, the more certain we can be about the performance of the Learning to Rank method. The methods for which it holds that there is no other method that has 1) a higher \ac{NWN} and 2) a higher number data sets evaluated, are identified as the best performing methods and are labelled with the name of the method.

\subsection{NDCG@3}
Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{NDCG}@3 metric. Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\

@@

LambdaNeuralRank and CoList both acquired a perfect \ac{NWN} score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2. LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}). FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in \ac{NWN}, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\

@@

LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a \ac{NWN} of 1.0. LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of data sets, but score decreasingly well in terms of \ac{NWN}. These results are highly in agreement with the \ac{NDCG}@3 comparison. The only modification compared to the \ac{NDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{NDCG}@5 comparison, but not in the \ac{NDCG}@3 comparison. Like in the \ac{NDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.

@@

LambdaMART and LambdaNeuralRank score a \ac{NWN} of 1.0 on the \ac{NDCG}@10 comparison. For LambdaNeuralRank these results are again based on AOL dataset measurements. LambdaMART showed the highest \ac{NDCG}@10 performance for the MSLR-WEB10k dataset. The set of algorithms for which there is no other algorithm with both a higher \ac{NWN} and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}. SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices. DCMP is not in the set in contrast with the \ac{NDCG}@3 and \ac{NDCG}@5 comparison.

\subsection{MAP}
Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning to Rank methods for the \ac{MAP} metric. Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw \ac{NWN} data for the Learning to Rank methods.\\

@@

Where comparisons on the \ac{NDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results. RankDE scores a \ac{NWN} of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{NDCG}-comparisons. In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\

LARF and LRUF score very high \ac{NWN} scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons. Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison. A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}. SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods. RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations. No reported \ac{NDCG} performance was found in the literature study for RE-QR. There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.

\subsection{Cross-metric}
Figure \ref{fig:normalised_winning_number_all} shows the \ac{NWN} as function of \ac{IWN} for the methods described in Table \ref{tab:ltr_methods_used}. Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.\\

@@

LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both \ac{IWN} and \ac{NWN}. LRUF is the only method that achieved this in all \ac{NDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison. With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{NDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{NDCG} performance as opposed to the \ac{MAP} performance. This was to be expected, because the cross-metric comparison input data of three \ac{NDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\

@@

C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect \ac{NWN} measures, but have low \ac{IWN} measures. Further evaluation runs of these methods on benchmark data sets that they are not yet evaluated on are desirable. The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more data sets than the number of data sets that we included evaluation results for in this meta-analysis. Some of the DirectRank measurements could not be used because measurements on some data sets were only available in graphical form and not in raw data.\\

@@

\section{Limitations}
In the \ac{NWN} calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark. By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set. It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some data sets than on other data sets. \\

@@

Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step. This brings up a risk of overly optimistic evaluation results affecting our NWN results. Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely. By taking IWN into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as IWN is loosely related with the number of studies that reported evaluation results for an algorithm.\\

Our comparison regarded evaluation results on \ac{NDCG}@{3, 5, 10} and \ac{MAP}. By making the decision to include \ac{NDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{NDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics. This implicit weighting could be regarded as arbitrary, but the number of algorithm evaluation results gained by this makes it a pragmatic approach. Note that another implicit weighting lies in the paper dimension. Hence, the higher number of evaluation results specified in a paper, the higher the influence of this paper on the outcome of the analysis. This implicit weighting is not harmful to the validity of our comparison, as papers with a large number of evaluation results are more valuable than papers with a few evaluation results. In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.

\section{Conclusions}
We proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets. Our comparison methodology comprises of two components: 1) \ac{NWN}, which provides insight in the ranking accuracy of the learning to rank method, and 2) \ac{IWN}, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\

Based on this this new comparison approach for a set of sparse evaluation results, we will now look back on the first research question of the thesis.
\begin{description}
\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
\end{description}
Although no closing arguments can be formulated on which Learning to Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\

Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others. However, no closing arguments can be formulated on which learning to rank methods are most accurate. LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy. From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\

@@

More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}. Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\

In the following chapters of this thesis, concerning parallel execution of the Learning to Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning to Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet. Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.\\

### 2014-09-29 report/Chapters/Future.tex
crossbenchmark comparison verbeterd vanuit IPM versie

\section{Computing models}
We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms. Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Spark \cite{}, Storm \cite{} and MPI \cite{}. Of these distributed programming models, Spark is particularly promising, since Shukla et al. \cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce.

>>>>

\section{Computing models}
We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms. Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and MPI. Of these distributed programming models, Spark is particularly promising, since Shukla et al. \cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce. For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice. This lack of a critical mass results in higher integration costs and less available support. The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem. Cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce.
